QOCO: A QoE-Oriented Computation Offloading Algorithm based on Deep Reinforcement Learning for Mobile Edge Computing

### 2023-10-15 paper-conference.tex
F_revision_2

We define two binary variables, $x_i(t)$ and $y_{i,j}(t)$ for $i \in \mathcal{I}$ and $j \in \mathcal{J}$ to determine the offloading decision and offloading target, respectively. Specifically, $x_i(t)$ indicates whether task $z_i(t)$ is assigned to the execution queue ($x_i(t) = 0$) or to the dispatch queue ($x_i(t) = 1$), and $y_{i,j}(t)$ indicates whether task $z_i(t)$ is offloaded to EN $j \in \mathcal{J}$. If the task is dispatched to EN $j$, we set $y_{i,j}(t) = 1$; otherwise, $y_{i,j}(t) = 0$.

\begin{figure}
	\captionsetup{name=Fig.}
	\centering
	\includegraphics[width=1\linewidth]{queue}
	\vspace*{-5mm}
	\caption{An illustration MD $i \in \mathcal{I}$ and EN $j \in \mathcal{J}$ in the MEC system.}
	\vspace*{-3mm}
	\label{fig1}
\end{figure}

@@

 We consider that the tasks in the dispatch queues are dispatched to the appropriate ENs via the MD wireless interface. We denote the transmission rate of MD $i$'s interface when communicating with EN $j \in \mathcal{J}$ at time $t$ as $r_{i,j}(t)$. At time slot $t \in \mathcal{T}$, if task $z_i(t)$ is assigned to the dispatch queue for computation offloading, we define $l_i^{\text{T}}(t) \in \mathcal{T}$ to represent the time slot when the task is either dispatched to the EN or dropped. We also define $\delta_i^{\text{T}}(t)$ as the number of time slots that task $z_i(t)$ should wait in the queue before transmission. It should be noted that MD $i$ computes the value of $\delta_i^{\text{T}}(t)$ before making a decision. The value of $\delta_i^{\text{T}}(t)$ is computed as follows:

@@

where $[\cdot]^+ =$ max$(0, \cdot)$ and $l_i^{\text{T}}(0)=0$ for the simplicity of presentation. Note that the value of $\delta_i^{\text{T}}(t)$ only depends on $l_i^{\text{T}}(t)$ for $t' < t$. If MD $i \in \mathcal{I}$ assigns task $z_i(t)$ for dispatching at time slot $t \in \mathcal{T}$, then it will either be dispatched or dropped at time slot $l_i^{\text{T}}(t)$, which is

@@

We model the local execution by a queuing system consisting the execution queue and the MD processor. Let $f_i$ denote the MD $i$'s processing power (in cycle per second). When task $z_i(t)$ is assigned to the execution queue at the beginning of time slot $t \in \mathcal{T}$, we define $l_i^{\text{C}}(t) \in \mathcal{T}$ as the time slot during which task $z_i(t)$ will either be processed or dropped. If the execution queue is empty, $l_i^{\text{C}}(t) = 0$. We also define $\delta_i^{\text{C}}(t)$ denote the number of remaining time slots before processing task $z_i(t)$ in the execution queue. The value of $\delta_i^{\text{C}}(t)$ is calculated  as follows:

@@

%In the above relation, the operator $[z]^+ = \max{0, z}$ is defined, and the variable $l_i^{\text{C}}(0) = 0$. Specifically, the expression $\max_{t' \in {0,1,2,\ldots,t-1}} l_i^{\text{C}}(t')$ determines the time slot until which all tasks placed in the local execution queue have been completed before time slot $t$. Therefore, $\delta_i^{\text{C}}(t)$ represents the number of time slots that task $z_i(t)$ needs to wait for processing. For example, suppose task $z_i(1)$ is in the execution queue, and its processing will be completed in time slot 5. Hence, $l_i^{\text{C}}(1) = 5$, meaning that the task should remain in the queue for $\Delta^{comp}(3) = [\max{5,0}-3+1]^+ = 3$ time slots until the process is finished.
In the equation above, the term $\max_{t' \in \{0, 1, \ldots, t-1\}} l_i^{\text{C}}(t')$ denotes the time slot at which each existing task in the execution queue, which arrived before time slot $t$, is either processed or dropped. Consequently, $\delta_i^{\text{C}}(t)$ denotes the number of time slots that task $z_i(t)$ should wait before being processed. We denote the time slot in which task $z_i(t)$ will be completely processed by $l_i^{\text{C}}(t)$ if it has been assigned to the execution queue for local processing in time slot $t$. We have

@@

@@

We model the edge execution by the queues associated with MDs deployed at ENs. If computation task $z_i(t')$ is dispatched to EN $j$ where $t' < t$, we let $z_{i,j}^{\text{E}}(t)$ and $\lambda_{i,j}^{\text{E}}(t)$ (in bits) denote the unique index of the task and the size of the task in the $m^{\text{th}}$ queue at EN $j$. We define $\eta_{i,j}^{\text{E}}(t)$ (in bits) as the length of this queue at the end of time slot $t \in \mathcal{T}$. We refer to a queue as an active queue in a certain time slot if it is not empty. That being said, if at least one task is already in the queue from previous time slots or there is a task arriving at the queue, that queue is active. We define $\mathcal{B}_j(t)$ to denote the active queue set at EN $j$ in time slot $t$, given by:

@@

We introduce $B_j(t) \triangleq |\mathcal{B}_j(t)|$ that represents the number of active queues in EN $j \in \mathcal{J}$ at time slot $t \in \mathcal{T}$. In each time slot $t \in \mathcal{T}$, the EN's processing power is divided among its active queues using a generalized processor sharing method~\cite{parekh1993generalized}. Let variable $f_j^{\text{E}}$ (in cycle per second) represent the computational capacity of EN $j$. Therefore, EN $j$ can allocate computational capacity of $f_j^{\text{E}}/(\rho_i(t) B_j(t))$ to each MD $i \in \mathcal{B}_j(t)$ during time slot $t$. To calculate the length of the execution queue for MD $i \in \mathcal{I}$ in EN $j \in \mathcal{J}$, we define $\omega_{i,j}(t)$ (in bits) to represent the number of bits from dropped tasks in that queue at the end of time slot $t \in \mathcal{T}$. The backlog of the queue, referred to as $\eta_{i,j}^{\text{E}}(t)$ is given by:

@@

Additionally, we define $D_{i,j}^{\text{E}}(t)$ to represent the quantity of processing time slots allocated to task $z_{i,j}^{\text{E}}(t)$ when executed at EN $j$. This value is given by:
\begin{alignat}{2}
	D_{i,j}^{\text{E}}(t) = { \lambda_{i,j}^{\text{E}}(t) \rho_i(t) \over f_j^{\text{E}} \tau /  B_j(t)}.
	\label{14}  
\end{alignat}
We define $E_{i,j}^{\text{E}}(t)$ as the energy consumption of processing at EN $j$ in time slot $t$ by MD $i$. This can be calculated as:
\begin{alignat}{2}
	E_{i,j}^{\text{E}}(t) = {D_{i,j}^{\text{E}}(t)  p_j^{\text{E}} \tau \over B_j(t)},  % = { \lambda_{i,j}^{\text{E}}(t) \rho_i(t) \over  f_j^{\text{E}}}p_j^{\text{E}} ,
	\label{15}  
\end{alignat}
where $p_j^{\text{E}}$ is a constant value which denotes the energy consumption of the EN $j$'s processor when operating at full capacity.

In addition to the energy consumed by the EN $j$ for task processing, we also take into account the energy consumed by the MD $i$'s user interface in the standby state while waiting for task completion at the EN $j$. We define $E_{i,j}^{\text{I}}(t)$ as the energy consumption associated with the user interface of MD $i \in \mathcal{I}$, which is given by
\begin{alignat}{2}
	E_i^{\text{I}}(t) = D_{i,j}^{\text{E}}(t) p_i^{\text{I}} \tau, %= {\lambda_{i,j}^{\text{E}}(t) \rho_i(t) \over \mathcal{B}_j(t) f_j^{\text{E}}} p_i^{\text{I}},
	\label{16}
\end{alignat}
where $p_i^{\text{I}}$ is the energy consumption of MD $i \in \mathcal{I}$ in the standby state. Accordingly, by combining \eqref{4}, \eqref{15} and \eqref{16}, the overall energy consumption associated with offloading task $z_i(t)$ is formulated as
\begin{alignat}{2}
	E_i^{\text{O}}(t) = E_i^{\text{T}}(t) + \sum_{\mathcal{J}} E_{i,j}^{\text{E}}(t) + E_i^{\text{I}}(t).
	\label{17}
\end{alignat}

@@

Based on the introduced system model, we present the computation task offloading problem in this section. Our primary goal is to enhance each MD's QoE individually by taking the dynamic demands of MDs into account. To achieve this, we approach the optimization problem as an MDP, aiming to maximize the MD's QoE by striking a balance among key QoE factors, including task completion, task delay, and energy consumption. To prioritize QoE factors, we utilize the MD's battery level, which plays a crucial role in decision-making. Specifically, when an MD observes its state (e.g. task size, queue details, and battery level) and encounters a newly arrived task, it is required to select an appropriate action for that task. The selected action, based on the observed state, will result in enhanced QoE for the MD. Each MD strives to maximize its long-term QoE by optimizing the policy mapping from states to actions. In what follows, we will initially present the state space, action space, and QoE function, respectively. Then, we will formulate the QoE maximization problem for each MD.

\subsection{State Space}
A state in our MDP represents a conceptual space that comprehensively describes the state of an MD facing the environment. We represent the MD $i$'s state at time slot $t$ as vector $\boldsymbol{s}_i(t)$ that includes the newly arrived task size, the queues information, the MD's battery level, and the workload history at the ENs. The MD observers this vector at the beginning of each time slot. The vector $\boldsymbol{s}_i(t)$ is defined as follows:
\begin{alignat}{2}
\boldsymbol{s}_i(t) = \Big(\lambda_i(t), \delta_i^{\text{C}}(t), \delta_i^{\text{T}}(t), \boldsymbol{\eta}_i^{\text{E}}\hspace{-0.4mm}(\hspace{-0.4mm}t\hspace{-0.4mm}-\hspace{-0.4mm}1\hspace{-0.5mm}),\phi_i(t), \mathcal{H}(t) \Big),
	\label{18}
\end{alignat}
where the vector $\boldsymbol{\eta}_i^{\text{E}}(t-1) = (\eta_{i,j}^{\text{E}}(t-1))_{j \in \mathcal{J}}$ represents the queue length of MD $i$ in ENs at the previous time slot, which the MD can compute this vector according to \eqref{10}. Let $\phi_i(t)$ denote the battery level of MD $i$ at time slot $t$. Considering the power modes of Android MDs, $\phi_i(t)$ is derived from the discrete set $\Phi=\{\phi_1,\phi_2,\phi_3\}$, corresponding to ultra power-saving, power-saving, and performance modes, respectively.

In addition, to predicting future EN workloads, we define the matrix $\mathcal{H}(t)$ as historical data, indicating the number of active queues for each EN. This data is recorded over $T^{\text{s}}$ time slots, ranging from $t-T^{\text{s}}$ to $t-1$, in a matrix with a size of $T^{\text{s}} \times J$. For EN $j$ workload history at $i^{th}$ time slot from $T^{\text{s}}-t$, define $h_{i,j}(t)$ as:
\begin{alignat}{2}
	h_{i,j}(t) = B_j(t - T^{\text{s}} + i - 1),
	\label{19}
\end{alignat}
%where $B_j(t - T^{\text{s}} + i - 1)$ is the number of active queues of EN $j$ in time slot $t - T^{\text{s}} + i - 1$. 
Where EN $j \in \mathcal{J}$ broadcasted $B_j(t)$ at the end of each time slot. Given $B_j(t) \leq I$, can be represented using only a few bits, the broadcast incurs minimal signaling overhead.

@@

\subsection{Action Space}
The action space represents the agent's behavior and the decisions made by the agent. In this context, we define $\boldsymbol{a}_i(t)$ to denote the action taken by MD $i \in \mathcal{I}$ at time slot $t \in \mathcal{T}$. These actions involve two decision, (a) Offloading decision to determine whether to offload the task, and (b) Offloading target to determine the EN to send the offloaded tasks. Thus, the action of MD $i$ in time slot $t$ can be concisely expressed as the following action tuple:
\begin{alignat}{2}
	\boldsymbol{a}_i(t) = (x_i(t), \boldsymbol{y}_i(t)),
	\label{20}
\end{alignat}
where the vector $\boldsymbol{y}_i(t)=(y_{i,j}(t))_{j \in \mathcal{J}}$ represents the selected EN for offloading this task. %Hence, the decision space of each MD is thus equal to $\mathcal{A} = \{0,1\}^{1+N}$, with $j$ representing the number of available ENs.

\subsection{QoE Function}
The QoE function evaluates the influence of an agent's actions by taking into account several key performance factors. Given the selected action $\boldsymbol{a}_i(t)$ in the observed state $\boldsymbol{s}_i(t)$, we represent $\mathcal{D}_i(\boldsymbol{s}_i(t), \boldsymbol{a}_i(t))$ as the delay of task $z_i(t),$ which indicates the number of time slots from time slot $t$ to the time slot in which task $z_i(t)$ is processed. It is calculated by:\\\\
$\mathcal{D}_i(\boldsymbol{s}_i(t),\boldsymbol{a}_i(t)) = (1-x_i(t))\Big(l_i^{\text{C}}(t) - t + 1\Big)  +$
\begin{alignat}{2}
	\;\;\;\; x_i(t)\bigg( \sum\limits_{\mathcal{J}} \sum\limits_{t'=t}^{T}\mathbbm{1}\big(z_{i,j}^{\text{E}}(t')=z_i(t)\big) l_{i,j}^{\text{E}}(t') - t +1\bigg),
	\label{21}  
\end{alignat} 
where $\mathcal{D}_i(\boldsymbol{s}_i(t),\boldsymbol{a}_i(t))= 0$ when task $z_i(t)$ is dropped. Correspondingly, we denote the energy consumption of task $z_i(t)$ with concerning the selected action $\boldsymbol{a}_i(t)$ in the observed state $\boldsymbol{s}_i(t)$ as $\mathcal{E}_i(\boldsymbol{s}_i(t),\boldsymbol{a}_i(t))$, which is:\\\\
$\mathcal{E}_i(\boldsymbol{s}_i(t),\boldsymbol{a}_i(t)) = (1-x_i(t)) E_i^{\text{L}}(t)+$
\begin{alignat}{2}
\hspace{1.8cm} x_i(t)\bigg( \sum\limits_{\mathcal{J}} \sum\limits_{t^{'}=t}^{T}\mathbbm{1}\big(z_{i,j}^{\text{E}}(t') = z_i(t)\big) E_i^{\text{O}}(t)  \bigg).
	\label{22}  
\end{alignat}
Given the delay and energy consumtion of task $z_i(t)$, we also define $\mathcal{C}_i(\boldsymbol{s}_i(t),\boldsymbol{a}_i(t))$ denote the assosiate cost of task $z_i(t)$ given the action $\boldsymbol{a}_i(t)$ in the state $\boldsymbol{s}_i(t)$ as follows: \\\\
$\mathcal{C}_i(\boldsymbol{s}_i(t),\boldsymbol{a}_i(t)) =$
\begin{alignat}{2}
\phi_i(t) \, \mathcal{D}_i(\boldsymbol{s}_i(t),\boldsymbol{a}_i(t)) +(1-\phi_i(t)) \, \mathcal{E}_i(\boldsymbol{s}_i(t),\boldsymbol{a}_i(t)),
	\label{23}  
\end{alignat}
where $\phi_i(t)$ represents the MD $i$'s battery level. When the MD is operating in performance mode, the primary focus is on minimizing task delays, which results in an increase in delay costs. On the other hand, when the MD switches to ultra power-saving mode, the main attention is directed toward reducing power consumption, leading to an increase in energy consumption cost.

Finally, we define $\boldsymbol{q}_i(\boldsymbol{s}_i(t),\boldsymbol{a}_i(t))$ as the QoE associated with task $z_i(t)$ given the selected action $\boldsymbol{a}_i(t)$ and the observed state $\boldsymbol{s}_i(t)$. The QoE function is defined as follows:\\\\
$\boldsymbol{q}_i(\boldsymbol{s}_i(t),\boldsymbol{a}_i(t)) =$
\begin{alignat}{2}
 \hspace*{8mm}
	\begin{cases} 
		\mathcal{R} - \mathcal{C}_i(\boldsymbol{s}_i(t),\boldsymbol{a}_i(t)) & \hspace*{-2mm} \text{if task $z_i(t)$ processed,} \\
		- \hspace*{0.8mm} \mathcal{E}_i(\boldsymbol{s}_i(t),\boldsymbol{a}_i(t)) &	\hspace*{-2mm} \text{if task $z_i(t)$ dropped,}
	\end{cases}
	\label{26}  
\end{alignat}
where $\mathcal{R} > 0$ represents a constant reward for task completion. If task $z_i(t) = 0$, then $\boldsymbol{q}_i(\boldsymbol{s}_i(t), \boldsymbol{a}_i(t)) = 0$. Throughout the rest of this paper, we adopt the shortened notation $\boldsymbol{q}_i(t)$ to represent $\boldsymbol{q}_i(\boldsymbol{s}_i(t), \boldsymbol{a}_i(t))$.
%When $\phi_i(t)$ is closer to 1, the cost function prioritizes reducing delay and drop, which could be advantageous in situations where tasks need to be completed quickly, and responsiveness is crucial. Conversely, when $\phi_i(t)$ is closer to 0, the cost function prioritizes reducing energy consumption, which is beneficial when the MD's battery level is low, and conserving power is necessary to prolong operation.

>>>>

We define two binary variables, $x_i(t)$ and $y_{i,j}(t)$ for $i \in \mathcal{I}$ and $j \in \mathcal{J}$ to determine the offloading decision and offloading target, respectively. Specifically, $x_i(t)$ indicates whether task $z_i(t)$ is assigned to the computation queue ($x_i(t) = 0$) or to the transmission queue ($x_i(t) = 1$), and $y_{i,j}(t)$ indicates whether task $z_i(t)$ is offloaded to EN $j \in \mathcal{J}$. If the task is dispatched to EN $j$, we set $y_{i,j}(t) = 1$; otherwise, $y_{i,j}(t) = 0$.

@@

 We consider that the tasks in the transmission queues are dispatched to the appropriate ENs via the MD wireless interface. We denote the transmission rate of MD $i$'s interface when communicating with EN $j \in \mathcal{J}$ in time $t$ as $r_{i,j}(t)$. In time slot $t \in \mathcal{T}$, if task $z_i(t)$ is assigned to the transmission queue for computation offloading, we define $l_i^{\text{T}}(t) \in \mathcal{T}$ to represent the time slot when the task is either dispatched to the EN or dropped. We also define $\delta_i^{\text{T}}(t)$ as the number of time slots that task $z_i(t)$ should wait in the queue before transmission. It should be noted that MD $i$ computes the value of $\delta_i^{\text{T}}(t)$ before making a decision. The value of $\delta_i^{\text{T}}(t)$ is computed as follows:

@@

where $[\cdot]^+ =$ max$(0, \cdot)$ and $l_i^{\text{T}}(0)=0$ for the simplicity of presentation. Note that the value of $\delta_i^{\text{T}}(t)$ only depends on $l_i^{\text{T}}(t)$ for $t' < t$. If MD $i \in \mathcal{I}$ schedules task $z_i(t)$ for dispatching at time slot $t \in \mathcal{T}$, then it will either be dispatched or dropped at time slot $l_i^{\text{T}}(t)$, which is

@@

We model the local execution by a queuing system consisting the computation queue and the MD processor. Let $f_i$ denote the MD $i$'s processing power (in cycle per second). When task $z_i(t)$ is assigned to the computation queue at the beginning of time slot $t \in \mathcal{T}$, we define $l_i^{\text{C}}(t) \in \mathcal{T}$ as the time slot during which task $z_i(t)$ will either be processed or dropped. If the computation queue is empty, $l_i^{\text{C}}(t) = 0$. Let $\delta_i^{\text{C}}(t)$ denote the number of remaining time slots before processing task $z_i(t)$ in the computation queue. The value of $\delta_i^{\text{C}}(t)$ is calculated  as follows:

@@

%In the above relation, the operator $[z]^+ = \max{0, z}$ is defined, and the variable $l_i^{\text{C}}(0) = 0$. Specifically, the expression $\max_{t' \in {0,1,2,\ldots,t-1}} l_i^{\text{C}}(t')$ determines the time slot until which all tasks placed in the local computation queue have been completed before time slot $t$. Therefore, $\delta_i^{\text{C}}(t)$ represents the number of time slots that task $z_i(t)$ needs to wait for processing. For example, suppose task $z_i(1)$ is in the computation queue, and its processing will be completed in time slot 5. Hence, $l_i^{\text{C}}(1) = 5$, meaning that the task should remain in the queue for $\Delta^{comp}(3) = [\max{5,0}-3+1]^+ = 3$ time slots until the process is finished.
In the equation above, the term $\max_{t' \in \{0, 1, \ldots, t-1\}} l_i^{\text{C}}(t')$ denotes the time slot at which each existing task in the computation queue, which arrived before time slot $t$, is either processed or dropped. Consequently, $\delta_i^{\text{C}}(t)$ denotes the number of time slots that task $z_i(t)$ should wait before being processed. We denote the time slot in which task $z_i(t)$ will be completely processed by $l_i^{\text{C}}(t)$ if it is assigned to the computation queue for local processing in time slot $t$. We have

@@

\begin{figure}
	\captionsetup{name=Fig.}
	\centering
	\includegraphics[width=1\linewidth]{queue}
	\vspace*{-5mm}
	\caption{An illustration of MD $i \in \mathcal{I}$ and EN $j \in \mathcal{J}$ in the MEC system.}
	\vspace*{-3mm}
	\label{fig1}
\end{figure}

@@

We model the edge execution by the queues associated with MDs deployed at ENs. If computation task $z_i(t')$ is dispatched to EN $j$ at time $t' < t$, we let $z_{i,j}^{\text{E}}(t)$ and $\lambda_{i,j}^{\text{E}}(t)$ (in bits) denote the unique index of the task and the size of the task in the $i^{\text{th}}$ queue at EN $j$. We define $\eta_{i,j}^{\text{E}}(t)$ (in bits) as the length of this queue at the end of time slot $t \in \mathcal{T}$. We refer to a queue as an active queue in a certain time slot if it is not empty. That being said, if at least one task is already in the queue from previous time slots or there is a task arriving at the queue, that queue is active. We define $\mathcal{B}_j(t)$ to denote the set of active queues at EN $j$ in time slot $t$, given by:

@@

We introduce $B_j(t) \triangleq |\mathcal{B}_j(t)|$ that represents the number of active queues in EN $j \in \mathcal{J}$ in time slot $t \in \mathcal{T}$. In each time slot $t \in \mathcal{T}$, the EN's processing power is divided among its active queues using a generalized processor sharing method~\cite{parekh1993generalized}. Let variable $f_j^{\text{E}}$ (in cycles per second) represent the computational capacity of EN $j$. Therefore, EN $j$ can allocate computational capacity of $f_j^{\text{E}}/(\rho_i(t) B_j(t))$ to each MD $i \in \mathcal{B}_j(t)$ during time slot $t$. To calculate the length of the computation queue for MD $i \in \mathcal{I}$ in EN $j \in \mathcal{J}$, we define $\omega_{i,j}(t)$ (in bits) to represent the number of bits from dropped tasks in that queue at the end of time slot $t \in \mathcal{T}$. The backlog of the queue, referred to as $\eta_{i,j}^{\text{E}}(t)$ is given by:

@@

Additionally, we define $D_{i,j}^{\text{E}}(t)$ to represent the quantity of processing time slots allocated to task $z_{i,j}^{\text{E}}(t)$ when executed at EN $j$. This value is given by:
\begin{alignat}{2}
	D_{i,j}^{\text{E}}(t) = { \lambda_{i,j}^{\text{E}}(t) \rho_i(t) \over f_j^{\text{E}} \tau /  B_j(t)}.
	\label{14}  
\end{alignat}
We also define $E_{i,j}^{\text{E}}(t)$ as the energy consumption of processing at EN $j$ in time slot $t$ by MD $i$. This can be calculated as:
\begin{alignat}{2}
	E_{i,j}^{\text{E}}(t) = {D_{i,j}^{\text{E}}(t)  p_j^{\text{E}} \tau \over B_j(t)},  % = { \lambda_{i,j}^{\text{E}}(t) \rho_i(t) \over  f_j^{\text{E}}}p_j^{\text{E}} ,
	\label{15}  
\end{alignat}
where $p_j^{\text{E}}$ is a constant value which denotes the energy consumption of the EN $j$'s processor when operating at full capacity.

In addition to the energy consumed by EN $j$ for task processing, we also take into account the energy consumed by the MD $i$'s user interface in the standby state while waiting for task completion at the EN $j$. We define $E_{i,j}^{\text{I}}(t)$ as the energy consumption associated with the user interface of MD $i \in \mathcal{I}$, which is given by
\begin{alignat}{2}
	E_i^{\text{I}}(t) = D_{i,j}^{\text{E}}(t) p_i^{\text{I}} \tau, %= {\lambda_{i,j}^{\text{E}}(t) \rho_i(t) \over \mathcal{B}_j(t) f_j^{\text{E}}} p_i^{\text{I}},
	\label{16}
\end{alignat}
where $p_i^{\text{I}}$ is the energy consumption of MD $i \in \mathcal{I}$ in the standby state. Accordingly, by combining \eqref{4}, \eqref{15} and \eqref{16}, the overall energy consumption associated with offloading task $z_i(t)$ is formulated as
\begin{alignat}{2}
	E_i^{\text{O}}(t) = E_i^{\text{T}}(t) + \sum_{\mathcal{J}} E_{i,j}^{\text{E}}(t) + E_i^{\text{I}}(t).
	\label{17}
\end{alignat}

@@

Based on the introduced system model, we present the computation task offloading problem in this section. Our primary goal is to enhance each MD's QoE individually by taking the dynamic demands of MDs into account. To achieve this, we approach the optimization problem as an MDP, aiming to maximize the MD's QoE by striking a balance among key QoE factors, including task completion, task delay, and energy consumption. To prioritize QoE factors, we utilize the MD's battery level, which plays a crucial role in decision-making. Specifically, when an MD observes its state (e.g. task size, queue details, and battery level) and encounters a newly arrived task, it selects an appropriate action for that task. The selected action, based on the observed state, will result in enhanced QoE. Each MD strives to maximize its long-term QoE by optimizing the policy mapping from states to actions. In what follows, we first present the state space, action space, and QoE function, respectively. We then formulate the QoE maximization problem for each MD.

\subsection{State Space}
A state in our MDP represents a conceptual space that comprehensively describes the state of an MD facing the environment. We represent the MD $i$'s state in time slot $t$ as vector $\boldsymbol{s}_i(t)$ that includes the newly arrived task size, the queues information, the MD's battery level, and the workload history at the ENs. The MD observers this vector at the beginning of each time slot. The vector $\boldsymbol{s}_i(t)$ is defined as follows:
\begin{alignat}{2}
\boldsymbol{s}_i(t) = \Big(\lambda_i(t), \delta_i^{\text{C}}(t), \delta_i^{\text{T}}(t), \boldsymbol{\eta}_i^{\text{E}}\hspace{-0.4mm}(\hspace{-0.4mm}t\hspace{-0.4mm}-\hspace{-0.4mm}1\hspace{-0.5mm}),\phi_i(t), \mathcal{H}(t) \Big),
	\label{18}
\end{alignat}
where vector $\boldsymbol{\eta}_i^{\text{E}}(t-1) = (\eta_{i,j}^{\text{E}}(t-1))_{j \in \mathcal{J}}$ represents the queues length of MD $i$ in ENs at the previous time slot and is computed by the MD according to \eqref{10}. Let $\phi_i(t)$ denote the battery level of MD $i$ in time slot $t$. Considering the power modes of a real mobile device, $\phi_i(t)$ is derived from the discrete set $\Phi=\{\phi_1,\phi_2,\phi_3\}$, corresponding to ultra power-saving, power-saving, and performance modes, respectively.

In addition, to predict future EN workloads, we define the matrix $\mathcal{H}(t)$ as historical data, indicating the number of active queues for all ENs. This data is recorded over $T^{\text{s}}$ time slots, ranging from $t-T^{\text{s}}$ to $t-1$, in a matrix with a size of $T^{\text{s}} \times J$. For EN $j$ workload history at $i^{th}$ time slot from $T^{\text{s}}-t$, we define $h_{i,j}(t)$ as:
\begin{alignat}{2}
	h_{i,j}(t) = B_j(t - T^{\text{s}} + i - 1).
	\label{19}
\end{alignat}
%where $B_j(t - T^{\text{s}} + i - 1)$ is the number of active queues of EN $j$ in time slot $t - T^{\text{s}} + i - 1$. 
EN $j \in \mathcal{J}$ broadcasts $B_j(t)$ at the end of each time slot. %Given $B_j(t) \leq I$, can be represented using only a few bits, the broadcast incurs minimal signaling overhead.

@@

\subsection{Action Space}
The action space represents the agent's behavior and the decisions. In this context, we define $\boldsymbol{a}_i(t)$ to denote the action taken by MD $i \in \mathcal{I}$ in time slot $t \in \mathcal{T}$. These actions involve two decisions, (a) Offloading decision to determine whether or not to offload the task, and (b) Offloading target to determine the EN to send the offloaded tasks. Thus, the action of MD $i$ in time slot $t$ can be concisely expressed as the following action tuple:
\begin{alignat}{2}
	\boldsymbol{a}_i(t) = (x_i(t), \boldsymbol{y}_i(t)),
	\label{20}
\end{alignat}
where vector $\boldsymbol{y}_i(t)=(y_{i,j}(t))_{j \in \mathcal{J}}$ represents the selected EN for offloading this task. %Hence, the decision space of each MD is thus equal to $\mathcal{A} = \{0,1\}^{1+N}$, with $j$ representing the number of available ENs.

\subsection{QoE Function}
The QoE function evaluates the influence of agent's actions by taking several key performance factors into account. Given the selected action $\boldsymbol{a}_i(t)$ in the observed state $\boldsymbol{s}_i(t)$, we represent $\mathcal{D}_i(\boldsymbol{s}_i(t), \boldsymbol{a}_i(t))$ as the delay of task $z_i(t),$ which indicates the number of time slots from time slot $t$ to the time slot in which task $z_i(t)$ is processed. It is calculated by:\\\\
$\mathcal{D}_i(\boldsymbol{s}_i(t),\boldsymbol{a}_i(t)) = (1-x_i(t))\Big(l_i^{\text{C}}(t) - t + 1\Big)  +$
\begin{alignat}{2}
	\;\;\;\; x_i(t)\bigg( \sum\limits_{\mathcal{J}} \sum\limits_{t'=t}^{T}\mathbbm{1}\big(z_{i,j}^{\text{E}}(t')=z_i(t)\big) l_{i,j}^{\text{E}}(t') - t +1\bigg),
	\label{21}  
\end{alignat} 
where $\mathcal{D}_i(\boldsymbol{s}_i(t),\boldsymbol{a}_i(t))= 0$ when task $z_i(t)$ is dropped. Correspondingly, we denote the energy consumption of task $z_i(t)$ when taking action $\boldsymbol{a}_i(t)$ in the observed state $\boldsymbol{s}_i(t)$ as $\mathcal{E}_i(\boldsymbol{s}_i(t),\boldsymbol{a}_i(t))$, which is:\\\\
$\mathcal{E}_i(\boldsymbol{s}_i(t),\boldsymbol{a}_i(t)) = (1-x_i(t)) E_i^{\text{L}}(t)+$
\begin{alignat}{2}
\hspace{1.8cm} x_i(t)\bigg( \sum\limits_{\mathcal{J}} \sum\limits_{t^{'}=t}^{T}\mathbbm{1}\big(z_{i,j}^{\text{E}}(t') = z_i(t)\big) E_i^{\text{O}}(t)  \bigg).
	\label{22}  
\end{alignat}
Given the delay and energy consumtion of task $z_i(t)$, we also define $\mathcal{C}_i(\boldsymbol{s}_i(t),\boldsymbol{a}_i(t))$ that denotes the assosiate cost of task $z_i(t)$ given the action $\boldsymbol{a}_i(t)$ in the state $\boldsymbol{s}_i(t)$. \\\\
$\mathcal{C}_i(\boldsymbol{s}_i(t),\boldsymbol{a}_i(t)) =$
\begin{alignat}{2}
\phi_i(t) \, \mathcal{D}_i(\boldsymbol{s}_i(t),\boldsymbol{a}_i(t)) +(1-\phi_i(t)) \, \mathcal{E}_i(\boldsymbol{s}_i(t),\boldsymbol{a}_i(t)),
	\label{23}  
\end{alignat}
where $\phi_i(t)$ represents the MD $i$'s battery level. When the MD is operating in performance mode, the primary focus is on minimizing task delays, thus the delay contributes more to the cost. On the other hand, when the MD switches to ultra power-saving mode, the main attention is directed toward reducing power consumption.

Finally, we define $\boldsymbol{q}_i(\boldsymbol{s}_i(t),\boldsymbol{a}_i(t))$ as the QoE associated with task $z_i(t)$ given the selected action $\boldsymbol{a}_i(t)$ and the observed state $\boldsymbol{s}_i(t)$. The QoE function is defined as follows:\\\\
$\boldsymbol{q}_i(\boldsymbol{s}_i(t),\boldsymbol{a}_i(t)) =$
\begin{alignat}{2}
 \hspace*{8mm}
	\begin{cases} 
		\mathcal{R} - \mathcal{C}_i(\boldsymbol{s}_i(t),\boldsymbol{a}_i(t)) & \hspace*{-2mm} \text{if task $z_i(t)$ is processed,} \\
		- \hspace*{0.8mm} \mathcal{E}_i(\boldsymbol{s}_i(t),\boldsymbol{a}_i(t)) &	\hspace*{-2mm} \text{if task $z_i(t)$ is dropped,}
	\end{cases}
	\label{26}  
\end{alignat}
where $\mathcal{R} > 0$ represents a constant reward for task completion. If $z_i(t) = 0$, then $\boldsymbol{q}_i(\boldsymbol{s}_i(t), \boldsymbol{a}_i(t)) = 0$. Throughout the rest of this paper, we adopt the shortened notation $\boldsymbol{q}_i(t)$ to represent $\boldsymbol{q}_i(\boldsymbol{s}_i(t), \boldsymbol{a}_i(t))$.
%When $\phi_i(t)$ is closer to 1, the cost function prioritizes reducing delay and drop, which could be advantageous in situations where tasks need to be completed quickly, and responsiveness is crucial. Conversely, when $\phi_i(t)$ is closer to 0, the cost function prioritizes reducing energy consumption, which is beneficial when the MD's battery level is low, and conserving power is necessary to prolong operation.
